---
title   : "**Black Friday Analysis**"
Authors : "Likky Sai Prakash, Siva Nagendra Babu"
output  : html_document
---
  
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


```{r,echo=FALSE}
library(data.table)
library(DataExplorer) # For initial exploratory data analysis
library(dplyr) # For data manipulation
library(ggplot2)
library(vcd)
library(rpart)
```

## **About our Data**:
* The dataset here is a sample of the transactions made in a retail store.
* The store wants to know better the customer purchase behaviour against different products.
* Specifically, here the problem is a regression problem where we are trying to predict the dependent variable (the amount of purchase) with the help of the information contained in the other variables.


* Things needed to be addressed:
  + Understanding the customers on the basis of their purchasing habits.
+ Understanding the purchasing habits according to Age groups, Occuptation, City_Categories.
+ The above segmented group of users can be then used to model the data and use to predict the purchase
spend for each customer.

* Lets dive in by understanding the data.


selecting Data 

```{r,echo=FALSE}
sales<-read.csv(file.choose())
```


# 1. Data Analysis

## **a. Let's see how are our actually data looks like**

```{r,echo=FALSE}
head(sales,10)
```

### **Observations**:
* 1.If we observe The Id column values are repeated with same numbers and it won't useful much in analysis
  * 2.Here age column data is distributed in interval range 
  * 3.There are total3 city categories
  * 4.There are 3 different product Categories available to customers

## **B. Lets see the shape of our data**

```{r, echo=FALSE}
dim(sales)
```

## **c. Let's get the Description of our Data**
  
  ```{r, echo=FALSE}
str(sales)
```

### **Observations**:
* We have Categorical and Discrete Data in our Data set
* If we observe the Martial Status is of of Int where it should be Categorical
* Regarding product Categories We observe some missing values, we need to Handle those.

## **D. Let's see the missing values present in our Data**

```{r, echo=FALSE}
sapply(sales, function(x) sum(is.na(x)))
plot_missing(sales)
```


### **Observations** :
* We have missing values in Product_Category_2 and Product_category_3.

## **E.Now we'll try to understand about variable distribution.**

* Continuous variables 

```{r, echo=FALSE}
plot_histogram(sales)
```

* Discrete variables

+ Gender variable    

```{r, echo=FALSE}
prop.table(table(sales$Gender))
```

+ Customer Age group

```{r, echo=FALSE}
prop.table(table(sales$Age))
```

+ Which city has highest customers

```{r, echo=FALSE}
prop.table(table(sales$City_Category))
```

+ Ration of category Stay_In_Current_City_Years

```{r, echo=FALSE}
prop.table(table(sales$Stay_In_Current_City_Years))
```

## **Lets Draw the above Data via graphical intution**
```{r, echo=FALSE}
plot_bar(sales)
```

# 2. Data Transformation 

Lets Transform Features of our Data their Respective Data Type

```{r,}
sales$User_ID <- as.factor(sales$User_ID)
sales$Product_ID <- as.factor(sales$Product_ID)
sales$Age <- as.factor(sales$Age)
sales$Occupation <- as.factor(sales$Occupation)
sales$City_Category <- as.factor(sales$City_Category)
sales$Stay_In_Current_City_Years <- as.factor(sales$Stay_In_Current_City_Years)
sales$Product_Category_1 <- as.integer(sales$Product_Category_1)
sales$Product_Category_2 <- as.integer(sales$Product_Category_2)
sales$Product_Category_3 <- as.integer(sales$Product_Category_3)
sales$Purchase <- as.numeric(sales$Purchase)
```



# 3.Exploratory Data Analysis:



## Let's See who have purchased more so lets draw a count plot for it


```{r,echo=FALSE}
ggplot(sales, aes(Gender, Purchase)) + 
  geom_bar( stat = "identity", fill = "steelblue", color = "blue") +
  ggtitle("Total male and female Purchased") +
  theme(legend.position = "none")
```

Male count is More compared to female count while black friday sales purchasing

## Similarly let's see who have purchased more married or Unmarried

```{r,echo=FALSE}
ggplot(sales, aes(Marital_Status, Purchase)) + 
  geom_bar( stat = "identity", fill = "steelblue", color = "blue") +
  ggtitle("Total male and female Purchased") +
  theme(legend.position = "none")
```

Unmarried have bought more than the married people

## **1.Age group/gender category vs product category**:

```{r, echo=FALSE}

ggplot(sales, aes(Gender, Product_Category_1, fill = Gender)) + geom_col(width = 0.4) + facet_wrap(~ Age) + 
  labs(title = "Age Group/Gender Vs Product Category 1")

ggplot(sales, aes(Gender, Product_Category_2, fill = Gender)) + geom_col(width = 0.4) + facet_wrap(~ Age) + 
  labs(title = "Age Group/Gender Vs Product Category 2")

ggplot(sales, aes(Gender, Product_Category_3, fill = Gender)) + geom_col(width = 0.4) + facet_wrap(~ Age) + 
  labs(title = "Age Group/Gender Vs Product Category 3")
```



### **Observations:**

* So if we observe whether its product_category 1,2,3 most people from the age group 26-35 have purchased the    most.
* And the next the people in age group of 18-25 and 36-45 have purchased most
* With no surprise the females are the one's who done heavy shopping compared to men which is obvious in the     respective age group. 

## **2. Occupation vs purchases**:

```{r,echo=FALSE}
ggplot(sales, aes(Occupation, Product_Category_1, fill = Occupation)) + geom_col() +  facet_wrap(~ Gender) +
  labs(title = "Occupation Vs Product Category 1")

ggplot(sales, aes(Occupation, Product_Category_2, fill = Occupation)) + geom_col() +  facet_wrap(~ Gender) +
  labs(title = "Occupation Vs Product Category 2")

ggplot(sales, aes(Occupation, Product_Category_3, fill = Occupation)) + geom_col() +  facet_wrap(~ Gender) +
  labs(title = "Occupation Vs Product Category 3")
```

* 

### **Observations:**

* People of job category from 0-7 have purchased most products

## **3. City Category vs Purchases:**

```{r, echo=FALSE}
ggplot(sales, aes(City_Category, Product_Category_1, fill = City_Category)) + geom_col() +  facet_wrap(~ Age) +
  labs(title = "City Category Vs Product_Category 1")  
ggplot(sales, aes(City_Category, Product_Category_2, fill = City_Category)) + geom_col() +  facet_wrap(~ Age) +
  labs(title = "City Category Vs Product_Category 2")  
ggplot(sales, aes(City_Category, Product_Category_3, fill = City_Category)) + geom_col() +  facet_wrap(~ Age) +
  labs(title = "City Category Vs Product_Category 3")  

```
### **Observations:**

* People From city B with age group 26-35 have purchased most 
* Then city A with the similar group of have done most purchases



## 4. Stay in cuntry vs product category and Purchase amount

```{r,echo=FALSE}
ggplot(sales, aes(Stay_In_Current_City_Years,Product_Category_1 , fill = Stay_In_Current_City_Years)) + geom_col() +  
  facet_wrap(~ Age) + labs(title = "Stay in current city Vs Product_Category 1") 

ggplot(sales, aes(Stay_In_Current_City_Years,Product_Category_2 , fill = Stay_In_Current_City_Years)) + geom_col() +  
  facet_wrap(~ Age) + labs(title = "Stay in current city Vs Product_Category 2")

ggplot(sales, aes(Stay_In_Current_City_Years,Product_Category_3 , fill = Stay_In_Current_City_Years)) + geom_col() +  
  facet_wrap(~ Age) + labs(title = "Stay in current city Vs Product_Category 3") 
```


### **Observations:**

* People Staying with the period of one year and with age group 26-35 have done most purchases


## 6. Product Category vs Martial Status

```{r,echo=FALSE}
ggplot(sales, aes(Marital_Status, Product_Category_1, fill = Marital_Status)) + geom_col() +  
  facet_wrap(~ City_Category) + labs(title = "Marital Status Vs Product_Category 1 (wrapped with city category)")  
ggplot(sales, aes(Marital_Status, Product_Category_2, fill = Marital_Status)) + geom_col() +  
  facet_wrap(~ City_Category) + labs(title = "Marital Status Vs Product_Category 1 (wrapped with city category)")  
ggplot(sales, aes(Marital_Status, Product_Category_3, fill = Marital_Status)) + geom_col() +  
  facet_wrap(~ City_Category) + labs(title = "Marital Status Vs Product_Category 1 (wrapped with city category)")  
```

### **Observations:**

* People from city A with martial Status of not married have purchased more


# Let's see if there are any outliers with respect to Occupation data so we'll draw boxplot for checking it

```{r,echo=FALSE}
qplot(data =  sales,x = Occupation ,y = Purchase ,geom = "boxplot")
```

See We have few outliers in some areas of the Occupation 



# Conclusion:
  + CITY CATEGORY B HAD THE MOST CUSTOMERS(42%), FOLLOWED BY C (31%), AND LASTLY A (27%)
  + PEOPLE WHO HAD STAYED IN THE CITY FOR 1 YEAR CONSTITUTED MAJORITY OF THE CUSTOMERS
  + SINGLES MADE UP 59 % WHILE MARRIED MADE UP 41% OF THE CUSTOMERS
  + AS EXPECTED 26-35 YEAR OLDS SPENT MORE MONEY
  + IN ALL THE CITY CATEGORIES, THE SINGLES SPENT THE MOST MONEY
  + FOR THE PRODUCT CATEGORIES WITH THE HIGHEST PURCHASE, THE PLOTS PRESENT THEM IN DETAILS
  


